---
title: "Aides Drug"
author: "Natalia Lekach"
date: "March 20, 2021"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
triUmp <-read.csv(file="triUmpsal.csv",header=TRUE)
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

A new  preventative aides drug (triUmpsal) is under test in clinical trials. Consider only the females, does this drug seem effective?  Hint - use dplyr to obtain that subpopulation with the filter command.

The null hypothesis is that there is no difference in infection rate among females who received the treatment and females who received no treatment.  
  
The alternative hypothesis is that there is a difference in infection rate among females who received the treatment and females who received no treatment.  

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->
  
To determine the effectiveness of this preventative AIDs drug on females, we will compare the infection rate among study participants who either received the drug treatment or did not. The female study participants were filtered out to properly respond to our question.    

```{r}
FemaleTriUmp<-filter(triUmp,gender=="F")
```
# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

```{r}
barchartGC(~infected + treatment,data=FemaleTriUmp)
barchartGC(~infected + treatment,data=FemaleTriUmp, type="percent")
```
  
Of the female patients diagnosed with AIDS, about 60% of them received the drug treatment and about 40% of them received no treatment. Of the patients that were not diagnosed with AIDS, a little less than half of them received the drug and a little more than half did not receive the drug.    

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
table1 <- xtabs(~infected + treatment, data = FemaleTriUmp)
rowPerc(table1)
colPerc(table1)
```
  
Of the female patients diagnosed with AIDS, more of them received the drugs than received the placebo. Of the female patients who were not diagnosed with AIDS, ~2% more of them received the placebo. It seems that the drug was not effective in preventing AIDS. The inferential test will help determine if there is a statistical difference.  
  
## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```
  
Due to the small amount of participants who were diagnosed with AIDS, the Fisher Exact Test is a better statistical test to interpret for this study. With a small p-value of 0.00006596, we can reject the null hypothesis. This indicates that there is a significant difference in AIDS infection rate between those those who receive the treatment and those that receive the placebo. 
The odds ratio tells us that those who received the treatment are 1.544 times more likely to get AIDS than those that received the placebo.   
  
# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->


